%---------------------------------------------------------------------
%
%                      abstract.tex
%
%---------------------------------------------------------------------
%
% Contiene el capítulo del resumen en inglés.
%
% Se crea como un capítulo sin numeración.
%
%---------------------------------------------------------------------

\chapter{Abstract}
\cabeceraEspecial{Abstract}

% \begin{FraseCelebre}
% \begin{Frase}
% Perfection is achieved, not when there is nothing more to add, 
% but when there is nothing left to take away.
% \end{Frase}
% \begin{Fuente}
% Antoine de Saint-Exupéry
% \end{Fuente}
% \end{FraseCelebre}


\section*{Game engine and editor 2D for non-programmers}

A game engine is a development environment that provides tools for the creation of video games.
tools for the creation of video games. These tools prevent the developer from
developer to implement a large amount of functionality in order to focus more on the
more on the development of the videogame. Some examples of functionalities provided by the engines are: graphic rendering, physics engine,
audio system, player input management, resources management, network management, etc.

In addition, an editor may be integrated. Editors are visual tools whose purpose is to communicate to the engine the actions performed by the developer. They are therefore part of the engine's development environment. Editors tend to have a slow learning curve, especially for those who are not familiar with the engine. especially for those who are not familiar with the engine in particular or with game development in general.
game development in general. However, once developers become familiar with the
Once developers become familiar with the tools, however, they can significantly speed up the process of game
the game creation process and improve productivity.

This is a great advantage to experienced developers, but engines like Unity or UnrealEngine
engines such as Unity or UnrealEngine can harbor too much complexity for non-programmers, even for
for people with no programming experience, even if they are aiming at simple 2D games.
simple 2D games. A very useful tool to solve this problem is visual programming.
problem is visual programming. This type of programming allows users to
users to create logic by manipulating graphical elements instead of specifying them exclusively textually.
rather than specifying them exclusively in textual form. Unity has
Unity Visual Scripting and UnrealEngine with Blueprints.

The engine of this graduate work consists of a self-sufficient 2D video game development environment. This means that it will allow
manage the video game resources, scenes and interactive elements. In addition, it will support the creation of behaviors through node-based visual
visual programming based on nodes, the execution of the game in the editor and the creation of nal
the creation of final executables of the game for distribution.

\endinput
% Variable local para emacs, para  que encuentre el fichero maestro de
% compilación y funcionen mejor algunas teclas rápidas de AucTeX
%%%
%%% Local Variables:
%%% mode: latex
%%% TeX-master: "../Tesis.tex"
%%% End:
